---
title: Generating Maximal Partial Room Squares in R
author: Matthew Henderson
output:
  github_document:
    toc: yes
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
references:
- id: meszkaMaximalPartialRoom2021
  abstract: >-
    A partial Room square is maximal if no further pair of elements can be
    placed into any unoccupied cell without violating the conditions that define
    a partial Room square. This article is concerned with determining the
    spectrum of volumes of maximal partial Room squares of order n where the
    volume is the number of occupied cells and the order n is even.
  accessed:
    - year: 2021
      month: 7
      day: 21
  author:
    - family: Meszka
      given: Mariusz
    - family: Rosa
      given: Alexander
  container-title: Journal of Combinatorial Designs
  DOI: 10.1002/jcd.21777
  ISSN: 1520-6610
  issue: '7'
  issued:
    - year: 2021
  language: en
  note: '_eprint: https://onlinelibrary.wiley.com/doi/pdf/10.1002/jcd.21777'
  page: 482-501
  source: Wiley Online Library
  title: Maximal partial Room squares
  type: article-journal
  URL: https://onlinelibrary.wiley.com/doi/abs/10.1002/jcd.21777
  volume: '29'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/"
)

sapply(list.files("R", full.names = TRUE), source)
```

Two different greedy procedures
for building maximal partial Room squares
in R
inspired by
@meszkaMaximalPartialRoom2021.

# I: greedy1

`greedy1`
visits all cells in order placing
into the next cell the first available
pair not violating the conditions
of being a partial Room square.

```{r greedy1_alg}
library(tictoc)

n <- 10

tic()
R <- greedy1(n)
toc()
```

```{r greedy1_example_plot, echo=FALSE}
targets::tar_read(greedy1_example_plot)
```

```{r greedy1_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

# II: greedy2

`greedy2`
iterates through all pairs in order
placing the next pair in the first available cell
not violating the conditions of being a partial
Room square.

```{r greedy2_alg}
n <- 10

tic()
R <- greedy2(n)
toc()
```

```{r greedy2_example_plot, echo=FALSE}
targets::tar_read(greedy2_example_plot)
```

```{r greedy_2_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

# Results

```{r results}
targets::tar_read(results)
```

# References