---
title: Generating Maximal Partial Room Squares in R
author: Matthew Henderson
output:
  github_document:
    toc: yes
    toc_depth: 1
  html_document:
    toc: yes
editor_options: 
  chunk_output_type: console
references:
- id: meszkaMaximalPartialRoom2021
  abstract: >-
    A partial Room square is maximal if no further pair of elements can be
    placed into any unoccupied cell without violating the conditions that define
    a partial Room square. This article is concerned with determining the
    spectrum of volumes of maximal partial Room squares of order n where the
    volume is the number of occupied cells and the order n is even.
  accessed:
    - year: 2021
      month: 7
      day: 21
  author:
    - family: Meszka
      given: Mariusz
    - family: Rosa
      given: Alexander
  container-title: Journal of Combinatorial Designs
  DOI: 10.1002/jcd.21777
  ISSN: 1520-6610
  issue: '7'
  issued:
    - year: 2021
  language: en
  note: '_eprint: https://onlinelibrary.wiley.com/doi/pdf/10.1002/jcd.21777'
  page: 482-501
  source: Wiley Online Library
  title: Maximal partial Room squares
  type: article-journal
  URL: https://onlinelibrary.wiley.com/doi/abs/10.1002/jcd.21777
  volume: '29'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "figure/"
)

sapply(list.files("R", full.names = TRUE), source)
```

Two different greedy procedures
for building maximal partial Room squares
in R
inspired by
@meszkaMaximalPartialRoom2021.

# I: greedy1

`greedy1`
visits all cells in order placing
into the next cell the first available
pair such that the conditions
of being a partial Room square
will not be violated.

```{r greedy1_example_plot}
R <- greedy1(6)
plot_room_square_labs(R)
```

In this plot colour represents the order in which cells were filled.
The cells were filled in order from lightest to darkest.

```{r greedy1_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

## A few examples of maximal partial Room squares constructed with greedy1

```{r greedy1_examples_1, echo=FALSE}
library(patchwork)

R1 <- greedy1(8)
R2 <- greedy1(10)

(plot_room_square_labs(R1) + plot_room_square_labs(R2))
```

```{r greedy1_examples_2, echo=FALSE}
R3 <- greedy1(12)

plot_room_square_labs(R3)
```

# II: greedy2

`greedy2`
iterates through all pairs in order
placing the next pair in the first available cell
such that the conditions of being a partial
Room square are not violated.

```{r greedy2_example_plot}
R <- greedy2(8)
plot_room_square_labs(R)
```

Again colour represents order of filling, proceeding from lightest to darkest.

```{r greedy_2_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

## A few examples of maximal partial Room squares constructed with greedy2

```{r greedy2_examples_1, echo=FALSE, fig.cap="Put a caption here"}
R1 <- greedy2(8)
R2 <- greedy2(10)

(plot_room_square_labs(R1) + plot_room_square_labs(R2))
```

```{r greedy2_examples_2, echo=FALSE}
R3 <- greedy2(12)

plot_room_square_labs(R3)
```

# References