---
output: html_document
editor_options: 
  chunk_output_type: console
---

`greedy4`
is a modification of
`greedy3`.

`greedy4`
visits cells in order
placing into the next cell
the first available cell from
$M(R_t) \cap S \backslash {S(e, R_t) \choose 2}$.

The difference is that instead
of calculating
$T(R_t, e) := M(R_t) \cap S \backslash {S(e, R_t) \choose 2}$
at each step
we keep track of it
as cells are filled.

This means initialising
$T(R_t, e) \leftarrow {S \choose 2}$
at the beginning
and then after filling an empty cell
$e$
with
$p = \{x, y\}$:

  * removing every pair $p\prime$ containing either $x$ or $y$ from
  all lists of available pairs $e\prime$ in the same row or column as $e$:
  $T(R_{t + 1}, e\prime) \leftarrow T(R_t, e\prime)\backslash p\prime$
  * removing $p$ from the list of available pairs for every
  other empty cell $e\prime$: $T(R_{t + 1}, e\prime) \leftarrow T(R_t, e\prime)\backslash p$

```{r greedy4_alg}
n <- 14

tic()
R <- greedy4(14)
toc()
```

```{r greedy4_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

```{r greedy4_plot, echo=FALSE, message=FALSE, fig.width=9, fig.height=9}
nf <- n_filled_cells(R)
v <- volume(R)

ggplot(data = R %>% pivot_wider(), aes(col, row)) +
  geom_segment(data = grid_lines(n - 1, n - 1), aes(x = x, y = y, xend = xend, yend = yend), linewidth = .1) +
  geom_text(data = R %>% pivot_wider() %>% filter(!is.na(first)), aes(label = paste(first, second, sep = ","))) +
  scale_y_reverse() +
  coord_fixed() + 
  theme_void() +
  theme(
    legend.position  = "none"
  ) +
  labs(
    title = glue("A maximal partial Room square of order {n}"),
    subtitle = glue("Number of filled cells: {nf}, Volume: {v}.")
  )
```
