---
output: html_document
editor_options: 
  chunk_output_type: console
---

`greedy3`
is a modification of
`greedy1`.

`greedy3`
visits
all cells in order placing into the next cell
the first pair
$p = \{x, y\}$
in
$M(R_t) \cap S \backslash {S(e, R_t) \choose 2}$.
In other words the
first pair that is compatible
with the empty cell
which hasn't already been used
in
$R_t$

After filling an empty cell the global set of available
pairs is updated:
$U(R_{t + 1}) \leftarrow U(R_t)\backslash p$

As are the sets of symbols
not seen by empty cells
$e\prime$
lying in the same row
or column
as $e$:
$S\backslash S(e\prime, R_{t + 1}) \leftarrow S(e, R) \backslash {x}$
and
$S\backslash S(e\prime, R_{t + 1}) \leftarrow S(e, R) \backslash {y}$.

```{r greedy3_alg}
n <- 10

tic()
R <- greedy3(n)
toc()
```

```{r greedy3_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

```{r greedy3_plot, echo=FALSE, message=FALSE, fig.width=9, fig.height=9}
nf <- n_filled_cells(R)
v <- volume(R)

ggplot(data = R %>% pivot_wider(), aes(col, row)) +
  geom_segment(data = grid_lines(n - 1, n - 1), aes(x = x, y = y, xend = xend, yend = yend), linewidth = .1) +
  geom_text(data = R %>% pivot_wider() %>% filter(!is.na(first)), aes(label = paste(first, second, sep = ","))) +
  scale_y_reverse() +
  coord_fixed() + 
  theme_void() +
  theme(
    legend.position  = "none"
  ) +
  labs(
    title = glue("A maximal partial Room square of order {n}"),
    subtitle = glue("Number of filled cells: {nf}, Volume: {v}.")
  )
```
