---
output: html_document
editor_options: 
  chunk_output_type: console
---

`greedy3`
is a modification of
`greedy1`.

`greedy3`
visits
all cells in order placing into the next cell
the first pair
$p = \{x, y\}$
in
$M(R_t) \cap S \backslash {S(e, R_t) \choose 2}$.
In other words the
first pair that is compatible
with the empty cell
which hasn't already been used
in
$R_t$

After filling an empty cell the global set of available
pairs is updated:
$U(R_{t + 1}) \leftarrow U(R_t)\backslash p$

As are the sets of symbols
not seen by empty cells
$e\prime$
lying in the same row
or column
as $e$:
$S\backslash S(e\prime, R_{t + 1}) \leftarrow S(e, R) \backslash {x}$
and
$S\backslash S(e\prime, R_{t + 1}) \leftarrow S(e, R) \backslash {y}$.


```{r greedy3_init_alt, include=FALSE}
# this is needed simply because when I use graph
# intersection i have to use the symbols 1..n
# instead of 0..n-1 because igraph
# doesn't like 0 for a vertex label
# seems like i ought to rewrite intersect_G to handle
# that case rather than rewrite this function
# or generalise it
not_used_pairs <- function(R) {
  n <- max(R$row) + 1
  # find the pairs already used
  used_pairs <- R %>% select(first, second)
  # construct the set of pairs not used (namely all pairs minus used pairs)
  x <- combn(1:n, 2)

  all_pairs <- tibble(
    first = as.numeric(x[1,]),
    second = as.numeric(x[2,])
  )

  anti_join(all_pairs, used_pairs, by = c("first", "second")) %>%
    mutate(ffs = map2(first, second, c)) %>%
    pull(ffs)
}
```

```{r greedy3_alg}
n <- 10

tic()
R <- expand_grid(row = 1:(n - 1), col = 1:(n - 1)) %>%
  mutate(first = as.numeric(NA), second = as.numeric(NA)) %>%
  mutate(avail = list(1:n)) %>%
  greedy3()
toc()
```

```{r greedy3_is_maximal_room}
# is R a maximal partial Room square?
is_maximal_proom(R)
```

```{r greedy3_plot, echo=FALSE, message=FALSE, fig.width=9, fig.height=9}
nf <- n_filled_cells(R)
v <- volume(R)

ggplot(data = R %>% pivot_wider(), aes(col, row)) +
  geom_segment(data = grid_lines(n-1, n-1), aes(x = x, y = y, xend = xend, yend = yend), size = .1) +
  geom_text(data = R %>% pivot_wider() %>% filter(!is.na(first)), aes(label = paste(first, second, sep = ","))) +
  scale_y_reverse() +
  coord_fixed() + 
  theme_void() +
  theme(
    legend.position  = "none"
  ) +
  labs(
    title = glue("A maximal partial Room square of order {n}"),
    subtitle = glue("Number of filled cells: {nf}, Volume: {v}.")
  )
```
